---
title: "ADAproject"
output: html_document
date: "2022-11-03"
---
# set up
```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(plyr)
library(dplyr)
library(haven)
library(VIM)
library(table1)
library(pacman)
library(odds.n.ends)
pacman::p_load(readr, nnet, MASS, funModeling, table1, mutlcomp)
```

## 1) Load dataset
```{r}
CHIS2017 <- read_sav("/Users/manyi-eyongayuk/Downloads/adult_2017_spss/ADULT.sav")
```

#2) Data Cleaning:
# a) Separate out the variables i want because this dataset is HUGE

```{r}
CHIS2017 <- CHIS2017 %>%
dplyr::select(c("AC100", "BMI_P", "SRSEX", "SRAGE_P1"))
```
# b) Let's check how much data is missing though

```{r}
md.pattern(CHIS2017, rotate.names=TRUE)
table(CHIS2017$AC100, useNA="always")
# two ways to look at it
# no missing data :)
```
# c) Let's filter out the 184 people that answered "-1= inapplicable"
```{r}
CHIS2017 <- filter(CHIS2017, AC100 > -1)
```

# d) Recode BMI to an ordinal scale
```{r}

CHIS2017 <- CHIS2017 %>%
  mutate(BMI_cat = case_when(BMI_P > 0 & BMI_P < 18.5 ~ 0,
                             BMI_P >= 18.5 & BMI_P < 25 ~ 1,
                             BMI_P >= 25 & BMI_P < 30 ~ 2,
                             BMI_P >= 30 ~ 3)) %>%
  mutate(BMI_cat = factor(BMI_cat, 0:3, c("Underweight", "Normal", "Overweight", "Obese"), ordered = T))

```
# e) Condense the age category because we really don't need that many
```{r}
CHIS2017 <- CHIS2017 %>%
  mutate(AGE_cat = case_when(SRAGE_P1 >= 18 & SRAGE_P1 < 35 ~ 0,
                             SRAGE_P1 >= 35 & SRAGE_P1 < 65 ~ 1,
                             SRAGE_P1 >= 65 ~ 2))
```

# 3) Analysis
# a) Execute an ordinal logistic regression with frequency of exercise in past 7 days  as the independent variable and BMI category as the dependent variable with age and gender as covariates. 
```{r echo = FALSE, eval = FALSE}
# unadjusted model
model1<-polr(BMI_cat ~ AC100, Hess=TRUE, data=CHIS2017) 
summary(model1)

# adjusted model
model2<-polr(BMI_cat ~ AGE_cat + SRAGE_P1 + AC100, Hess=TRUE, data=CHIS2017) 
summary(model2)
```

# b) Get p-values
```{r}
ptable <- coef(summary(model1))
# calculate and store p values
p <- pnorm(abs(ptable[, "t value"]), lower.tail = FALSE) * 2
p
# combined table
ptable <- cbind(ptable, "p value" = p)
ptable
# The effects of single marital status on stage at diagnosis was statistically significant with a p-value <0.001 using married as the reference category.
```
#c) Get odds ratios
```{r}
exp(cbind("Odds ratio" = coef(model1), confint.default(model1, level = 0.95)))
exp(cbind("Odds ratio" = coef(model2), confint.default(model2, level = 0.95)))
```

# Interpretations
## For every one unit increase in frequency of exercise, the odds of a being obese decreases by 0.91 times after adjusting for sex and age


# 4) Testing the parallel regression/proportional odds assumption using Brant Test
```{r}
brant(model1)
brant(model2)

# The models do not satisfy the proportional odds assumption.
```
# 5) Switch to a multinomial logisitic regression :(
#a) First, go back and make underweight and normal one category to make this easier for me to interpret.
```{r}
CHIS2017 <- CHIS2017 %>%
  mutate(BMI = case_when(BMI_P > 0 & BMI_P < 25 ~ 0,
                             BMI_P >= 25 & BMI_P < 30 ~ 1,
                             BMI_P >= 30 ~ 2),
         BMI = factor(BMI, 0:2, c("Underweight/Normal", "Overweight", "Obese")))
```
# b) Okay, now let's re-run the model
```{r}
#unadjusted
model3 <- multinom(BMI ~ AC100, data=CHIS2017)
summary(model3)
tidy(model3, conf.int=TRUE, exponentiate = TRUE)

#adjusted
model4 <- multinom(BMI ~ AGE_cat + SRSEX +AC100, data=CHIS2017)
summary(model4)
tidy(model4, conf.int=TRUE, exponentiate = TRUE)


```

#6) DAG code
```{r}
testImplications <- function( covariance.matrix, sample.size ){
	library(ggm)
	tst <- function(i){ pcor.test( pcor(i,covariance.matrix), length(i)-2, sample.size )$pvalue }
tos <- function(i){ paste(i,collapse=" ") }
implications <- list(c("age","sex"))
	data.frame( implication=unlist(lapply(implications,tos)),
		pvalue=unlist( lapply( implications, tst ) ) )

}
```


# 7) Table 1
```{r}
#Turn gender and age category into factors
CHIS2017$SRSEX<-factor(CHIS2017$SRSEX,
														levels = c(1, 2),
														labels = c("Male", "Female"))
CHIS2017$AGE_cat<-factor(CHIS2017$AGE_cat,
														levels = c(0, 1, 2),
														labels = c("18-34", "35-64", "65+"))

table1(~AC100 + SRSEX + AGE_cat|BMI, CHIS2017)
# add labels
label(CHIS2017$SRSEX)<-"SELF-REPORTED SEX"
label(CHIS2017$AGE_cat)<-"AGE CATEGORY"
table1(~AC100 + SRSEX + AGE_cat|BMI, CHIS2017)

```


